---
title: "LDFA Vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{LDFA Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(LDFA)
```

# Assumptions and limitations

* 1) Unequal sample sizes are not a problem; however, the sample size within groups must be greater than the number of independent variable predictors (Rule of thumb: ~ 4 to 5 as many observations as predictor variables).

* 2) Outliers can cause problems when assessing significance and should transformed or removed. Normally distributed variables are preferred, but non-normally distributed variables will not influence model significance as long as the non-normality isn't driven by outliers.

* 3) Independent variables must uncorrelated and have low multicollinearity. 

* 4) There shouldn't be significant heterogeneity in the variance and co-variance matricies between groups. 

# How does a discriminant function anaysis work?

> The analysis creates "discriminant functions" using linear combinations of predictor variables. The number of functions is equal to either one less than the number of groups or the number of predictors—whichever is smaller. Each discriminant function maximizes the amount of discrimination between groups while also remaining non-correlated with the other discriminant functions. The model made from these discriminant functions can then be used to predict group membership on a new, uncharacterized sample.
# We use the built in dataset "iris" for this example.

> Split the data into training (80%) and test (20%) sets
```{r}
Partition(iris) 
x<- Partition(iris) # This function partitions your dataset into test and training subsets at a ratio of 20% to 80% as the default. You should store this in a separate variable which will keep two lists associated witht the test and training data.
```

> Create your model:
```{r}
DFA_pred(Species~., x$train)
y <- DFA_pred(Species~., x$train) # This function takes a specified model (your groups ~ predictor variables) and then uses that model to predict group membership of you specified data. In this case we have specified the training data.
z <- DFA_pred(Species~., x$test)
```

> Plotting the result in ggplot:
```{r}
DFA_plot(y)
g <- DFA_plot(y) # This function takes the prediction data derived from the DFA_pred function and plots it using ggplot.
g
```

> Now, test the accuracy of your model by calculating the fraction of correct classifications.
```{r}
Accuracy(y, NULL, x)
a <- Accuracy(y, z, x) # This function takes your predicted group assignments of the training and test data and compares it to the actual group assignments. It then takes the mean to show the porpotion correct for each dataset. In this case we did not use the test data to make predictions so we set the input to NULL. 
```
